---
title: "Quality traits prediction models using NIRS data"
author: "Chafik Analy"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load necessary  packages

```{r}
library(dplyr)
library(mdatools)
library(sf)
library(icardaFIGSr)

```

## Source API functions 

```{r}
## Please make sure to adjust the nir_api.R file path to use the API call

base::source("/Volumes/Macintosh HD — Data/Desktop/FIGS/icardaFIGSr/nir_api.R")

```

## Get Bread wheat Data

```{r}

## set parameters
Year = 2017:2020
QualityLab = "ICARDA-MAR"
Crop = "Bread wheat"


## Nir 
Nirdata <- getNIRData(qualityLab = QualityLab , crop = Crop,
                      year = Year )
## Traits 
Triatsdata <- getTraitsData(qualityLab = QualityLab , crop = Crop,
                      year = Year)

```

## Model Calibration using multiple smoothing and preprocessing methods.

```{r}
## gather trait and corresponding NIR data
#bw_spectra%>%count(Year)

# Subset Morocco and 2019
Train_test_data <- Nirdata%>%
  filter(Year==2019)%>%
  filter(Country=="Morocco")%>%
  left_join(Triatsdata, by = "QualityLabPlotNumber")%>%
  select(Protein, `400`:`2498`)%>%
  filter(!is.na(Protein))


# apply SG filter for smoothing
pspectra_svg = prep.savgol(Train_test_data[,-1], width = 15, porder = 3)

# apply SG filter for smoothing and take a first derivative
pspectra_1stD = prep.savgol(Train_test_data[,-1], width = 15, porder = 3, dorder = 1)

# apply SG filter for smoothing and take a second derivative
pspectra_2ndD = prep.savgol(Train_test_data[,-1], width = 15, porder = 3, dorder = 2)

# apply SNV 
pspectra_snv = prep.snv(Train_test_data[,-1])

# apply MSC 
pspectra_msc = prep.msc(as.matrix(Train_test_data[,-1]))


#unit area normalization
pspectra_normarea = prep.norm(Train_test_data[,-1], "area")
# unit length normalization
pspectra_normunit = prep.norm(Train_test_data[,-1], "length")

# show results
mdaplot(Train_test_data[,-1], type = "l", main = "Original")
mdaplot(pspectra_svg, type = "l", main = "SVG")
mdaplot(pspectra_snv, type = "l", main = "SNV")
mdaplot(pspectra_msc, type = "l", main = "MSC")
mdaplot(pspectra_normarea, type = "l", main = "Normal Area")
mdaplot(pspectra_normunit, type = "l", main = "Normal unit")
mdaplot(pspectra_1stD, type = "l", main = "SVG 1st Derivative")
mdaplot(pspectra_2ndD, type = "l", main = "SVG 2nd Derivative")
```

## train/test split 

### Train models on unprocessed Nir Data

```{r}
## Data partition index
trainIndex = seq(4, 720,3)

Xc <- Train_test_data[trainIndex, -1] # Calibration predictors
yc <- Train_test_data[trainIndex, 1]  # Calibration response

Xt <- Train_test_data[-trainIndex, -1] # Test predictors
yt <- Train_test_data[-trainIndex, 1]  # Test response


## Fit model on original data
model_raw <- pls(Xc, yc, 10, scale = TRUE,
         x.test = Xt, y.test = yt,
        info = "Protein prediction model")

```

## Train models on processed NIR data

### SNV

```{r}

Xc_snv <- pspectra_snv[trainIndex, -1] # Calibration predictors
yc_snv <- Train_test_data[trainIndex, 1]  # Calibration response

Xt_snv <- pspectra_snv[-trainIndex, -1] # Test predictors
yt_snv <- Train_test_data[-trainIndex, 1]  # Test response

model_snv <- pls(Xc_snv, yc_snv, 10, scale = TRUE, 
                 x.test = Xt_snv, y.test=yt_snv, 
        info = paste0("Protein prediction model with snv prep"))
```

### MSC 

```{r}

Xc_msc <- pspectra_msc[trainIndex, -1] # Calibration predictors
yc_msc <- Train_test_data[trainIndex, 1]  # Calibration response

Xt_msc <- pspectra_msc[-trainIndex, -1] # Test predictors
yt_msc <- Train_test_data[-trainIndex, 1]  # Test response

model_msc <- pls(Xc_msc, yc_msc, 10, scale = TRUE, 
                 x.test = Xt_msc, y.test=yt_msc, 
        info = paste0("Protein prediction model with msc prep"))
```

### SVG 

```{r}

Xc_svg <- pspectra_svg[trainIndex, -1] # Calibration predictors
yc_svg <- Train_test_data[trainIndex, 1]  # Calibration response

Xt_svg <- pspectra_svg[-trainIndex, -1] # Test predictors
yt_svg <- Train_test_data[-trainIndex, 1]  # Test response


model_svg <- pls(Xc_svg, yc_svg, 10, scale = TRUE, 
                 x.test = Xt_svg, y.test=yt_svg, 
        info = paste0("Protein prediction model with svg prep"))
```

### SVG 1st Derivative

```{r}

Xc_svg_1D <- pspectra_1stD[trainIndex, -1] # Calibration predictors
yc_svg_1D <- Train_test_data[trainIndex, 1]  # Calibration response

Xt_svg_1D <- pspectra_1stD[-trainIndex, -1] # Test predictors
yt_svg_1D <- Train_test_data[-trainIndex, 1]  # Test response


model_svg_1D <- pls(Xc_svg_1D, yc_svg_1D, 10, scale = TRUE, 
                 x.test = Xt_svg_1D, y.test=yt_svg_1D, 
        info = paste0("Protein prediction model with svg 1st Derivative prep"))
```

### SVG 2nd Derivative

```{r}

Xc_svg_2D <- pspectra_2ndD[trainIndex, -1] # Calibration predictors
yc_svg_2D <- Train_test_data[trainIndex, 1]  # Calibration response

Xt_svg_2D <- pspectra_2ndD[-trainIndex, -1] # Test predictors
yt_svg_2D <- Train_test_data[-trainIndex, 1]  # Test response



model_svg_2D <- pls(Xc_svg_2D, yc_svg_2D, 10, scale = TRUE, 
                 x.test = Xt_svg_2D, y.test=yt_svg_2D, 
        info = paste0("Protein prediction model with svg 2nd Derivative prep"))
```


## Get Models summary

```{r}
summary(model_raw)
summary(model_snv)
summary(model_svg)
summary(model_msc)
summary(model_svg_1D)
summary(model_svg_2D)
```

## Visualize models results 


```{r}
plot(model_raw)
```


```{r}
plot(model_snv)
```

```{r}
plot(model_msc)

```

```{r}
plot(model_svg)
```

```{r}
plot(model_svg_1D)
```

```{r}
plot(model_svg_2D)
```

## Plot model coefficients

```{r}
plotRegcoeffs(model_raw)
plotRegcoeffs(model_msc, ncomp = 5)
plotRegcoeffs(model_snv, ncomp = 5, type = "l", show.labels = F)
plotRegcoeffs(model_svg, ncomp = 5)
```

## plot models predictions

```{r}
plotPredictions(model_msc)
plotPredictions(model_snv)
plotPredictions(model_svg)

```

### Variables importance (wavelengths vs protein)

```{r}
plotVIPScores(model_msc)
plotVIPScores(model_msc, ncomp = 2, type = "h", show.labels = F)
plotSelectivityRatio(model_msc)
plotSelectivityRatio(model_msc, ncomp = 2, type = "h", show.labels = F)
```


```{r}

## Trying cross validated and autoscaled model on msc prep

vip <- vipscores(model_msc, ncomp = 2)


msc_scale_cv_model <- pls(Xc_msc, yc_msc, 10, scale = TRUE, 
           #      x.test = Xt_msc, y.test=yt_msc,
           center = TRUE, 
           cv = 10,
           exclcols = (vip <0.5),
        info = paste0("Protein prediction model with MSC prep"))



plotRegcoeffs(msc_scale_cv_model,
              ncomp = 2, type = "l", show.ci = TRUE,
              show.labels = F)
plotRegcoeffs(msc_scale_cv_model, ncomp = 3,
              type = "l", show.ci = TRUE, show.labels = F)
plotRegcoeffs(msc_scale_cv_model,ncomp = 4, type = "l",
              show.ci = TRUE,
              show.labels = F)
plotRegcoeffs(msc_scale_cv_model, ncomp = 5, type = "l",
              show.ci = TRUE, show.labels = F)

```

### Interval pls modeling

```{r}

msc_ipls_model = ipls(as.matrix(Xc_msc),as.matrix(yc_msc),
         scale = T, 
         glob.ncomp = 10, int.num = 10, full = FALSE)

summary(msc_ipls_model)
```

```{r}
plotRMSE(msc_ipls_model)
```

```{r}

plot(msc_ipls_model)
```

### Model prediction on test set

```{r}
plotPredictions(model_msc$res$test, show.stat = TRUE)
plotPredictions(model_snv$res$test, show.stat = TRUE)
plotPredictions(model_svg$res$test, show.stat = TRUE)
plotPredictions(model_svg_1D$res$test, show.stat = TRUE)
plotPredictions(model_svg_2D$res$test, show.stat = TRUE)

```

